An FMI-Based Initialization Plugin for INTO-CPS Maestro 2

### 2020-06-30 background.tex
Updates from Overleaf

\begin{definition}[Algebraic loops] 
The set of interconnected variables in an algebraic loop is defined as: $LoopVariables = \{v | v \in V : Path(v,v)\}$ \\
An algebraic loop is defined as the set $\{p_1, p_2 | p_1, p_2 \in LoopVariables: Path(p_1, p_2) \land Path(p_2, p_1)\}$\todo{Look at this - can it be specified better?}\\
Path is defined as the transitive closure of the set of relations: $\bigcup_{c \in \fmus} \set{\feedthrough{c}} \cup \coupling$\\
The algebraic dependency will lead to the following difference between performing to following :
\todo{Claudio check this definition and add a formalization of the challenge introduced}
\end{definition}

@@

As described by definition \ref{def:challenge} the algebraic loops introduce certain challenges in the system. This means that algebraic loops needs to be handled using fixed point iterations\cite{Gomes2018}. It is a technique to repeatedly perform the steps of a sub-list of the co-simulation step. The number of repetitions the operations should be be performed depends on the characteristics of the scenario. The operations should be performed until the system converge. 
An example of applying a fixed point iteration can be seen in Figure \ref{} where a bigger system contains an algebraic loop.

\begin{figure}[H]
    \centering
    \includegraphics{images/fmu_cycle.pdf}
    \caption{Co-simulation scenario with cycle}
    \label{fig:fmu_cycle}
\end{figure}

The initialization of the system requires application of fixed point iteration as defined by definition \ref{def:fixedpoint}.

\begin{definition}[Fixed point iteration of an Initialization procedure]\label{def:fixedpoint}
Given an Initialization procedure $\sequence{\initcall_i}_{i \in \setnat}$ of a co-simulation $\tuple{\fmus, \coupling}$ where $LoopVariables \neq \emptyset$.\\
Fixed point iteration should be used by 
$\forall v \in LoopVariables $

@@

\begin{definition}[Convergence of Fixed point iteration]\label{def:convergence}
A fixed point iteration is converging if a finite number of iterations will make the difference of the output-value of the same operation between two following iterations within a certain threshold.\\
$Convergence = \exists n \in \setnat: \forall p \in LoopVariables \implies |f(p, n + 1) - f(p, n)| \leq T$\\
$f$ is the operation performed on a loop-variable p of a given iteration n.
\end{definition}

>>>>

\begin{definition}[Algebraic loops] 
The set of interconnected variables in an algebraic loop is defined as: $\LoopVariables = \{v | v \in V : Path(v,v)\}$ \\
An algebraic loop is defined as the set $\{p_1, p_2 | p_1, p_2 \in \LoopVariables: Path(p_1, p_2) \land Path(p_2, p_1)\}$\todo{Look at this - can it be specified better?}\\
Path is defined as the transitive closure of the set of relations: $\bigcup_{c \in \fmus} \set{\feedthrough{c}} \cup \coupling$\\
The algebraic dependency will lead to the following difference between performing to following :
\todo{Claudio check this definition and add a formalization of the challenge introduced}
\end{definition}

@@

As described by definition \ref{def:challenge} the algebraic loops introduce certain challenges in the system. This means that algebraic loops needs to be handled using fixed point iterations\cite{Gomes2018}. It is a technique to repeatedly perform the steps of a sub-list of the co-simulation step. The number of repetitions the operations should be be performed depends on the characteristics of the scenario. The operations should be performed until the system converge. 
An example of applying a fixed point iteration can be seen in Figure \ref{def:fixedpoint} where a system containing an algebraic loop has to be initialized using fixed point iteration.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{images/fixedpoint.pdf}
    \caption{Fixed point iteration in a co-simulation scenario}
    \label{fig:fixedpont}
\end{figure}

The initialization of the system requires application of fixed point iteration as defined by definition \ref{def:fixedpoint}.
\begin{definition}[Fixed point iteration of an Initialization procedure]\label{def:fixedpoint}
Given an Initialization procedure $\sequence{\initcall_i}_{i \in \setnat}$ of a co-simulation $\tuple{\fmus, \coupling}$ where $\LoopVariables \neq \emptyset$.\\
Fixed point iteration should be used by 
$\forall v \in \LoopVariables $

@@

\begin{definition}[Convergence of Fixed point iteration]\label{def:convergence}
A fixed point iteration is converging if a finite number of iterations will make the difference of the output-value of the same operation between two following iterations within a certain threshold.\\
$Convergence = \exists n \in \setnat: \forall p \in \LoopVariables \implies |f(p, n + 1) - f(p, n)| \leq T$\\
$f$ is the operation performed on a loop-variable p of a given iteration n.
\end{definition}

### 2020-06-30 realization.tex
Updates from Overleaf

\subsection{Verification of the calculated Initialization Order}
The realized plugin has been verified in several different ways. Gomes et al. have in \cite{gomes_lucio_vangheluwe_2019} proved the correctness of using the topological order of a dependency graph of the interconnected variables to generate not only the initialization order but also the order of operations for a co-simulation step in master-algorithm. The purpose of the publication by Gomes et al. is a little different from the one of this paper. Gomes et al. are interested in both generating an initialization and master-algorithm, meaning they also need to focus on the time of each FMU. They also define their graph based on FMU-operations (\textit{Set, Get, doStep}) instead of interconnected variables, which is the approach described in this paper. This simplification is valid since this approach only cares about initialization and, therefore, does not need to care about the time of each FMU. This does that we can omit all the \textit{doStep} nodes and all edges connected to them from the graph by Gomes et al., this approach does, therefore, end up with a subgraph of the graph by Gomes et al. The proof from \cite{gomes_lucio_vangheluwe_2019} can therefore trivial be modified to the graph/approach presented in this paper.

Gomes et al.'s main contribution is an implementation of the FMI specification in Prolog freely available to be used in other projects. The Prolog implementation was used in their research to verify their proposed approach for the generation of co-simulation algorithms. The realization encapsulates all the rules of a valid master-algorithm and an initialization algorithm. As a part of this work, an integration from this plugin has been realized to interact with the Prolog realization. This integration has been realized in order to verify the correctness of the calculated initialization order against an established and recognized method. The integration is querying the Prolog database with a co-simulation environment and an initialization order. The integration requires a transformation of the dependency graph (see definition \ref{def:initialization_graph}) used in the plugin to a graph of FMU operations. This transformation is based on the definitions \ref{def:getout} and \ref{def:setin}. The described verification integration is not utilized if the graph contains algebraic loops since the Prolog database does not support these scenarios.

The Prolog code can generate an initialization order, but this functionality has not been used for several reasons. The main reason for not generating the initialization order using the Prolog code is based on performance considerations. The approach taken in this paper does also identifies the strongly connected components and performs a particular initialization strategy of these.

>>>>

\subsection{Verification of the Initialization Order}
The realized plugin has been verified in several ways using traditional proof methods and practical verification against an established co-simulation Verifier. The approach behind the plugin has been verified by Gomes et al. that in \cite{gomes_lucio_vangheluwe_2019} proved the correctness of using the topological order of a dependency graph of the interconnected variables to generate for the order of the operations that need to be performed in a co-simulation step (both the initialization procedure and an arbitrary step). Gomes et al. used a graph based on FMU-operations (\textit{Set, Get, doStep}) instead of interconnected variables, which is the approach described in this paper. The simplification of using the interconnected variables is valid since this approach only cares about initialization and, therefore, does not need to care about the time of each FMU. This does that we can omit all the \textit{doStep} nodes, it means that each interconnected variable can only perform one of the two operations \textit{get} or \textit{set} as described by the definitions \ref{def:getout} and \ref{def:setin}.
This approach does, therefore, end up with a subgraph of the graph by Gomes et al. The proof from \cite{gomes_lucio_vangheluwe_2019} can trivially be modified to the graph/approach presented in this paper.

\subsubsection{Practical verification against a Verifier} 
Gomes et al.'s main contribution is an implementation of the criteria for a valid FMI based co-simulation in Prolog freely available to be used in other projects\footnote{\url{}}. The Prolog implementation was used in their research to formally verify their proposed approach for the generation of co-simulation algorithms. The realization encapsulates all the rules of a valid master-algorithm and an initialization algorithm. As a part of this work, an integration from the developed plugin has been realized to interact with the Prolog Verifier. This integration has been realized to verify the correctness of the calculated initialization order against an established and recognized co-simulation Algorithm Generator and Verifier. The integration works by querying the Prolog database with a co-simulation environment and a calculated initialization to verify the order's correctness. The integration requires a transformation of the dependency graph (see definition \ref{def:initialization_graph}) used in the plugin to a graph of FMU operations. This transformation is based on the definitions \ref{def:getout} and \ref{def:setin}. The described verification integration is not utilized if the graph contains algebraic loops since the Prolog database does not support these scenarios.

The Prolog code can generate an initialization order, but this functionality has not been used for several reasons. The main reason for not creating the initialization order using the Prolog code is based on performance considerations. The approach taken in this paper does also identifies the strongly connected components and performs a particular initialization strategy of these.
